---
title: "Psychology Paper Analysis"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

# Summary of the paper

link to the article --> https://journals.sagepub.com/doi/pdf/10.1177/0956797616631990 

This article discusses choice making in humans stating that it is mostly attributed it to a top-down executive-control system. The top-down approach begins at the generally level where a problem is specified in general terms without details and is then broken down into smaller sub- problems. This process continues until the problems become small enough that one is able to solve them directly. However, living in an uncertain environment, action control may be more of an effect of both top-down and bottom-up factors. What was being explored in this study was whether separation of action and expectancy could occur in situations that are entirely predictable. This dissociation between expectancy and action would mean that despite knowing what will happen on the next trial, performance will still be affected by the bottom-up factors. What was examined in this study was the top-down control and that was done through the use of a go/no-go task whose trials alternated in an expected manner. The hypothesis was that an increase in motor excitability and a decrease in response latencies would be observed when a go response was expected by the subjects, the opposite would be true for an expected no-go response.

This study was conducted with sixteen students who were all University students; the subjects’ age range was 18-24 years. These subjects were asked to take part in a predictable go/no-go task. The subjects were asked which trial (go or no-go) they were expecting and their muscular responses to transcranial magnetic stimulation of the primary motor cortex was measured. An additional measure was that of response latencies, which was only done for the go trials. An iMac with Psychophysics Toolbox on was used to run this experiment. Go and no-go trials were presented to the subjects in runs of five, alternating predictably. Before beginning the trials, the motor threshold was determined. The trial began with a blank screen being shown for 3 to 4s. A numbering system was then used to ask the subjects how likely it was that a no-go stimulus would appear (1=not at all, 5=don’t know, and 9=definitely no-go trial). After the subjects answered this, the go or no-go stimulus appeared on the screen. Subjects had to respond, by pressing the mouse button, upon seeing the go stimulus and not respond when a no-go stimulus appeared. For half of them the go stimulus was the words “peanut butter” and the no-go stimulus was “brown sugar,” the opposite was true for the other half of the subjects. In order to indicate a mistake when a subject responded to a no-go stimulus, an error sound was played. This experiment had a total of 208 trials, which were split into 8 blocks and each block contained 28 trials. During the trials, MEP electrodes were attached to the subjects. There were two different times at which the TMS pulses were delivered. Pulse 1 was delivered during the blank interval and Pulse 2 was delivered when a rating was made by the subjects.

For data analysis, the expectancy ratings made by subjects during Go-1 trials were compared with the expectancy ratings for No-Go-5 trials and Go-2 trials, same for the no-go trials. The null hypothesis for this was there being no difference between trial types. Another set of reported data is the repeated measures analysis of variance with run type and run position. For the MEPs data, an exclusion criterion was determined which excluded any data that had shifted over 7mm away from what was the defined motor hot spot. This data was compared through the use of t-tests and Bayes factors for both Pulse 1 and Pulse 2. The numerical patterns that were observed for Pulse 1 and Pulse 2 trials were very similar. Two-way ANOVAs were run on this data in order to fully see the effect that the run type and run position had on MEPs. For the reaction times data, an exclusion was made of any reaction time (RT) that went above 2 standard deviations from the individual subject’s mean RT. After being averaged, this data was analyzed as a function of run position. This data was only for the go trials. Another data for the reaction times was a two-factor ANOVA using the variables pulse and run position. This indicated if
there was any effect on the RT data due to the pulse. Based on the ANOVA, a main effect of pulse was not seen and there was no interaction between the pulse and run position.

Based on Figure 2, it can be seen that through the trials the subjects were keeping track of the run sequence. Expectancy ratings were adjusted by the subjects after every run. There was a small difference seen within-run (i.e., Go 1-Go 2 and No Go1-No Go 2) when compared to the differences that were seen between-run. Figure 2 also shows that the MEP pattern had no correlation with the run sequence. The meta-analytic Bayes factors done showed large within-run differences, with an increase observed after the first go trial and a decrease after the first no-go trial. This overall showed that change occurred in the MEP between and within runs, with the differences being larger for within-run. Trial types, however, had no MEP difference. As for the expectancy analyses there were large between-run differences compared to the within-run differences, opposite with the MEP analyses. Staying consistent with the MEP data, the RT showed difference between the trials. Run of no-go trials resulted in longer RTs compared to the run of a single go trial.

The experimenters expected there to be correlation between the expectancy ratings, MEPs, and RTs as there could be a bias of actions by the top-down control system. As discussed, the expectancy ratings did not correspond with the MEP and RT data. There was no change in motor excitability observed between runs but there was a change after the first trial of a run, this showed that the results of the previous trial affected the motor excitability. RTs decreased as go trials went on. Based on the differences observed in the MEP data, it cannot be concluded that the changes are not due to the “retrieval of associations between the cue or stimulus and the go/no-go response.” The results, however, match the neural plasticity accounts. The main conclusion showed that, even in predictable environments, bottom-up influences modulate actions. Also, in situations where people are able to predict what’s going to happen next, people might still not adjust their behavior based on that prediction since it’s effortful.

# Re-analysis

```{r}
library(data.table)
df <- fread("~/Desktop/psyc 3400/paper analysis/ExpectTMS/R scripts/expectancy.csv")
```

```{r}
library(reshape)
library(ez)
data.wide <- read.csv('~/Desktop/psyc 3400/paper analysis/ExpectTMS/R scripts/expectancy.csv')

data.melt <- melt(data.wide, id.var="subject", variable_name="labels")
data.melt$labels <- as.character(data.melt$labels)
data.melt$type <- substr(data.melt$labels, 1,2)
data.melt$run <- substr(data.melt$labels, 3,3)
data.long <- data.frame(subject=data.melt$subject, type=data.melt$type, run=data.melt$run, exp=data.melt$value)

data.long$exp <- 10-data.long$exp
ezANOVA(
  data = data.long 
  , dv = .(exp)
  , wid = .(subject)
  , within = .(type, run)
  , detailed = T
)

source("pairwise_comparisons.R") 
TES(10-data.wide$no5, 10-data.wide$go1, paired = TRUE)
TES(10-data.wide$go1, 10-data.wide$go2, paired = TRUE)
TES(10-data.wide$go5, 10-data.wide$no1, paired = TRUE)
TES(10-data.wide$no1, 10-data.wide$no2, paired = TRUE)
```

```{r}
library(dplyr)


mean_accuracies <- data.long %>%
                    group_by(type,run) %>%
                    summarise(mean_accuracies = mean(exp))
```





Within-subjects ANOVA code for the effect of run position and run type on expectancy ratings
```{r}
S_accuracies <- data.long %>%
                    mutate(subject = as.factor(subject),
                           type = as.factor(type),
                           run  = as.factor(run)) %>%
                    group_by(subject,type,run) %>%
                    summarise(mean_accuracies = mean(exp))

summary(aov(mean_accuracies ~ type*run + Error(subject/(type*run)), S_accuracies))
```


# Re-analysis discussion

There were many ANOVA tests done in this study but the one the re-analysis was done for is the within-subjects ANOVA that checked the effect of run type and run position on expectancy ratings. The independent variables in this test were run type and run position which is why, in the code, the data is grouped by using these and the subjects. In the article, the values that were obtained upon doing the ANOVA analysis are provided on Table A2. The sum of the squares effect was 13.48, the sum of squares error was 60.95, and the F value was 3.31. Since the values I obtained through the ANOVA match those listed in the paper, it can be confirmed that I was able to reproduce the original analysis from the paper. The F value I obtained through the reanalysis was 3.318 and the other two values are exactly the same as listed previously. However, the one thing that is different is the p value. In the paper the p value is 0.077 while I obtained a p value of 0.016. Having a p value less than 0.05 means that the null hypothesis can be rejected and the effect is significant.  
